---
title: "A List of Codes for Econometric Analysis"
author: "Shunkei Kakimoto"
output:
  html_document:
    number_sections: yes
    # theme: flatly
    toc_float: yes
    toc: yes
    toc_depth: 2
geometry: margin=1in
---

```{r setup, include=FALSE}
library(knitr)
# opts_knit$set(root.dir = "")

# library(here)
# opts_knit$set(root.dir = here())

knitr::opts_chunk$set(
  echo = TRUE,
  cache = FALSE,
  comment = NA,
  message = FALSE,
  warning = FALSE,
  tidy = FALSE,
  cache.lazy = FALSE,
  #--- figure ---#
  dpi = 400,
  fig.width = 7.5,
  fig.height = 5,
  out.width = "750px",
  out.height = "500px"
)

# /*===== Basic Packages  =====*/
# /*---- Data Wrangling ----*/
library(data.table)
library(tidyverse)
library(DescTools)
library(maps)

# /*---- Visualization ----*/
library(RColorBrewer)
library(patchwork)
library(ggplot2)
library(ggthemes)
library(ggpubr)
library(viridis)
library(grid)
library(gridExtra)
library(GGally)

# --- OLS --- #
library(fixest)

# /*---- Model Summary ----*/
library(stats)
library(modelsummary)
library(flextable)
library(officer)
library(officedown)
library(gt)
```

# Motivation {- .unlisted}
+ Through this course, we will go through many R functions for practical econometric analysis (e.g. R functions for hypothesis testing, cluster-robust standard error estimation, panel data estimation methods, IV estimation method, ...). I thought it would be helpful to put all R functions introduced in this class in the same file so you don't need to search for the functions every time. This will be helpful especially when you conduct econometric analysis for your final report.

+ I'm going to update this material as we progress through the class.
+ Feel free to reach out if you want me to explain how to use a specific function. I will add more explanations accordingly. 


# Basic Regression Analysis: `feols` and `mlsummary`

```{r}
# === Preparation === #
library(modelsummary)
library(fixest)
library(dplyr)

# === Data === #
# install.packages("wooldridge") # if you have not installed yet
library(wooldridge)
# --- data "wage1" is loaded from R package "wooldridge" --- #
data("wage1") 
```

## Run simple regression {-}
+ Use `feols` function from the `fixest` package. 

```{r}
log_wage_model <- 
  feols(
    lwage ~ educ + exper + tenure,
    data = wage1
  )
```

### Quick Regression Summary {-}
In order to take a look at the regression results, the simplest way is to just evaluate the estimated regression model (`log_wage_model`)

```{r}
# === Quick summary === #
log_wage_model
```

## More on Regression Results {-}

+ The estimated regression modes (`log_wage_model`) contains the following information, and you can access them using `$`.

```{r}
# === The information contained in the "feols" output  === #
ls(log_wage_model)
```

```{r}
# === Coefficient of educ=== #
log_wage_model$coefficients[["educ"]]

# === Extract fitted values for sample points === #
# NOTE: show only the first five values
log_wage_model$fitted.values[1:5]
```

+ Prediction with the estimated regression model can be done with `predict` function.

```{r}
# NOTE: Change only the variable of educ. Fix the other variables using mean values.
newdata  <-  
  data.frame(
    educ = c(5, 10, 15),
    exper = mean(wage1$exper),
    tenure = mean(wage1$tenure)
  )

predict(log_wage_model,  newdata  =  newdata)
```


## Present Regression Results {-}
+ Use `msummary` function from the `modelsummary` package. The basic usage is the following.

```{r}
# === Preparation === #
library(modelsummary) 

# === report the summary=== #
msummary(
	log_wage_model, 
  title = 'This is the title',
	stars  =  TRUE,
	gof_omit  =  "IC|Log|Adj|F|Pseudo|Within" 
	)
```

+ You can present multiple regression results in the same table. 

```{r}
ls_res <- 
  list(
    "OLS 1" = feols(lwage ~ educ + exper + tenure, data = wage1),
    "OLS 2" = feols(lwage ~ educ + exper + expersq +tenure, data = wage1)
  )

msummary(
  ls_res,
  stars  =  TRUE, 
  gof_omit  =  "IC|Log|Adj|F|Pseudo|Within" 
  )
```

<br>

# Statistical  Hypothesis  Testing

## Get the CI of coefficients {-}

+ Below you find 90% CI (i.e., `conf.low` and `conf.high`) of coefficients of an example regression model  `log_wage_model`. 

```{r}
tidy(log_wage_model,  conf.int  =  TRUE,  conf.level  =  0.9)  %>% 
  relocate(term,  conf.low,  conf.high) %>%
  print()
```

## F-test {-}
+ Use `linearHypothesis` function from the `car` package

```{r, eval=FALSE}
# === Syntax === #
linearHypothesis(regression, hypothesis)
```

+ **NOTE**: To use `linearHypothesis()`, you need to use `lm()` to run regression instead to use `feols()`.



```{r}
# --- Load Package --- #
library(car)

# --- Load Data:mlb1 from "wooldridge" package --- #
data(mlb1)

#--- run a rergession ---#
mlb_reg <- 
  lm(
  log(salary) ~ years + gamesyr + bavg  + hrunsyr + rbisyr, 
  data = mlb1
  )

#--- See the regression results ---#
msummary(
  mlb_reg,
  stars  =  TRUE, 
  gof_omit  =  "IC|Log|Adj|F|Pseudo|Within" 
  )
```

Let's see the statistical significance of the joint effect of `bavg`, `hrunsyr`, and `rbisyr` on `salary`.

```{r}
# --- F-test --- #
linearHypothesis(mlb_reg, c("bavg=0", "hrunsyr=0", "rbisyr=0"))
```

<br>

# Standard Error Estimation

## Report Heteroskedasticity-robust SE estimates {-}
+ Add the `vcov` option like below inside `feols()` 

```{r}
# === Run Regression === #
reg_mlb <- 
  feols(
    log(salary) ~ years + bavg, 
    # --- assumes that errors are heteroskedastic --- # 
    vcov = "hetero", 
    data = mlb1
  ) 

# === Report the result as usual === #
msummary(
  reg_mlb,
  stars = TRUE,  
  gof_omit = "IC|Log|Adj|F|Pseudo|Within"
  )
```

## Report Cluster-robust SE estimates {-}
+ Cluster-robust variance covariance estimator take care of heteroskedasticity at the same time (so, they really are cluster- and heteroskedasticity-robust standard error estimators)
+ Add the `cluster` option like below inside `feols()` 

```{r}
# === Run Regression === #
reg_mlb  <-  
  feols(log(salary)  ~  years  +  bavg,  
  # --- assumes that errors are correlated within the cluster groups (nl)--- #
  cluster  =  ~  nl,  
  data  =  mlb1
  ) 

# === Report the result as usual === #
msummary(
  reg_mlb,
  stars = TRUE,  
  gof_omit = "IC|Log|Adj|F|Pseudo|Within"
  )
```



